\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-04-24 doc/USER_MANUAL/02_getting_started.tex
updates user manual to use facilitate the generation of a markdown version to the github wiki

% !TEX encoding = IsoLatin
%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english]{book}
\usepackage[T1]{fontenc}
%% DK DK we do not accept accented letters, use the old LaTeX syntax
%% DK DK for them instead: \'e \`e  \"e \^e and so on.
%% DK DK \usepackage[latin1]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{babel}
\usepackage{float}
\usepackage{textcomp}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},backref=section,colorlinks=false,pdfpagemode=None]
 {hyperref}
\hypersetup{
 pdftex,pdfstartview=FitH,pdfpagelayout=SinglePage}
\usepackage{breakurl}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\newcommand{\noun}[1]{\textsc{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxcode}
{\par\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\raggedright
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}%
 \item[]}
{\end{list}}
\newenvironment{lyxlist}[1]
{\begin{list}{}
{\settowidth{\labelwidth}{#1}
 \setlength{\leftmargin}{\labelwidth}
 \addtolength{\leftmargin}{\labelsep}
 \renewcommand{\makelabel}[1]{##1\hfil}}}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\renewcommand{\baselinestretch}{1.5}

% figures
\usepackage[dvips]{epsfig}

% fonts
\usepackage{times}

% hyperlinks to sections and references

\newcommand{\urlwithparentheses}[1]{(\url{#1})}

% biblio GJI
\bibliographystyle{abbrvnat}

\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
\newcommand{\eg}{e.g.,}

% colors for the text
\@ifundefined{definecolor}
 {\usepackage{color}}{}
\newcommand{\red}[1]{\textcolor{Red}{#1}}

% some of Dimitri's aliases for math symbols
\newcommand{\Grad}{\mbox{\boldmath $\bf \nabla$}}
\newcommand{\Div}{\Grad \cdot }
\newcommand{\Laplacian}{\nabla^2}
\newcommand{\pressure}{p}
\newcommand{\Int}{\displaystyle\int}
\newcommand{\Sum}{\displaystyle\sum}
\renewcommand{\epsilon}{\varepsilon}
\newcommand{\myiomega}{\mathbf{i} \, \omega}
\newcommand{\p}{\partial}
\newcommand{\boldu}{\mathbf{u}}
\newcommand{\boldv}{\mathbf{v}}
\newcommand{\boldw}{\mathbf{w}}
\newcommand{\boldx}{\mathbf{x}}
\newcommand{\boldy}{\mathbf{y}}
\newcommand{\boldm}{\mathbf{m}}
\newcommand{\boldp}{\mathbf{p}}
\newcommand{\boldg}{\mathbf{g}}
\newcommand{\bolds}{\mathbf{s}}
\newcommand{\boldn}{\mathbf{\hat{n}}}
\newcommand{\boldd}{\mathbf{d}}
\newcommand{\boldf}{\mathbf{f}}
\newcommand{\boldB}{\mathbf{B}}
\newcommand{\boldc}{\mathbf{c}}
\newcommand{\boldC}{\mathbf{C}}
\newcommand{\boldE}{\mathbf{E}}
\newcommand{\boldH}{\mathbf{H}}
\newcommand{\boldI}{\mathbf{I}}
\newcommand{\boldsigma}{\mbox{\boldmath $\bf \sigma$}}
\newcommand{\boldxi}{\mbox{\boldmath $\bf \xi$}}
\newcommand{\boldvarepsilon}{\mbox{\boldmath $\bf \varepsilon$}}
\newcommand{\bolddelta}{\mbox{\boldmath $\bf \delta$}}
\newcommand{\boldtau}{\mbox{\boldmath $\bf \tau$}}
\newcommand{\Boxmath}{\Lambda}
\newcommand{\rmd}{\, \mathrm{d}}
\newcommand{\pv}{\ ,}
\newcommand{\pp}{\ .}

\usepackage{babel}

\makeatother

\begin{document}
\begin{center}
\thispagestyle{empty}\vspace*{-1.8truecm} %
\makebox[1\textwidth]{%
\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover} %
}
\par\end{center}

\title{\textbf{SPECFEM3D Cartesian}\\
 \textbf{User Manual}}

\author{$\copyright$ Princeton University (USA), CNRS and University of Marseille (France),\\
and ETH Z\"urich (Switzerland)\\
Version 3.0}

\date{\today}

\maketitle

\section*{Authors}

\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
other equations. We are deeply indebted and thankful to him for that.
That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
We are deeply indebted and thankful to them for that.

%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
Since then it has been developed and maintained by a development team:
in alphabetical order,
Jean-Paul (Pablo) Ampuero,
Kangchen Bai,
Piero Basini,
C\'eline Blitz,
Alexis Bottero,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Paul Cristini,
Percy Galvez,
Dominik G\"oddeke,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Matthieu Lefebvre,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Surendra Nadh Somala,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Elliott Sales de Andrade,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Jean-Pierre Vilotte,
Zhinan Xie,
Chang-Hua Zhang,
Hejun Zhu.\\

The cover graphic of the manual was created by Santiago Lombeyda from
Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}.\\

The code is released open-source under the GNU version 2+ license, see the license at the end of this manual.\\

\newpage{}
\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.125\textwidth]{figures/logo_cnrs}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_princeton}\vspace*{2truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_ETH}\vspace*{2truemm}
\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_inria}\vspace*{2truemm}
\includegraphics[width=0.14\textwidth]{figures/logo_UPPA}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_IUF}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Caltech}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Harvard}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_IPGP}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_ANR}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_NSF}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.190\textwidth]{figures/logo_European_Union}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_GENCI}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_PRACE}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_CINES}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge}\vspace*{2truemm}
\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca}
\includegraphics[width=0.130\textwidth]{figures/logo_CIG}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto}\vspace*{2truemm}
\includegraphics[width=0.109\textwidth]{figures/logo_INGV}\vspace*{2truemm}
\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks}\vspace*{2truemm}
\includegraphics[width=0.120\textwidth]{figures/logo_CINECA}\vspace*{2truemm}
\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation}
\par\end{centering}
\end{figure}

\newpage{}

\tableofcontents{}

\chapter{Introduction}

The software package SPECFEM3D Cartesian simulates seismic wave propagation
at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin
technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
with optimized efficiency because of its tensorized basis functions
\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
very distorted mesh elements \citep{OlSe11}.\\

It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
The spectral element approach admits spectral rates of convergence
and allows exploiting $hp$-convergence schemes. It is also very well
suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
Tensor products inside each element can be optimized to reach very
high efficiency \citep{DeFiMu02}, and mesh point and element numbering
can be optimized to reduce processor cache misses and improve cache
reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
as well as mixed meshes, although with increased cost and reduced
accuracy in these elements, as in the discontinuous Galerkin method.\\

Note that in many geological models in the context of seismic wave
propagation studies (except for instance for fault dynamic rupture
studies, in which very high frequencies or supershear rupture need
to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
a continuous formulation is sufficient because material property contrasts
are not drastic and thus conforming mesh doubling bricks can efficiently
handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\

For a detailed introduction to the SEM as applied to regional seismic
wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate full 21-parameter anisotropy
(see~\citet{ChTr07}) as well as lateral variations in attenuation
\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\

The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave
propagation. Early seismic wave propagation applications of the SEM,
utilizing Legendre basis functions and a perfectly diagonal mass matrix,
include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
Chebyshev basis functions and a nondiagonal mass matrix include \citet{SePr94},
\citet{PrCaSe94} and \citet{SePrPr95}.\\

All SPECFEM3D software is written in Fortran2003 with full portability
in mind, and conforms strictly to the Fortran2003 standard. It uses
no obsolete or obsolescent features of Fortran. The package uses parallel
programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
supercomputing award in 2010.\\

It reached the sustained one petaflop performance level for the first time in February 2013
on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\

This new release of the code includes Convolution or
Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\

The next release of the code will use the PT-SCOTCH parallel and threaded
version of SCOTCH for mesh partitioning instead of the serial version.\\

SPECFEM3D Cartesian includes coupled fluid-solid domains and adjoint
capabilities, which enables one to address seismological inverse problems,
but for linear rheologies only so far. To accommodate visco-plastic
or non-linear rheologies, the reader can refer to the \href{http://geoelse.stru.polimi.it/}{GeoELSE}software
package \citep{CaGa97,StPaIg09}.

\section{Citation}

If you use SPECFEM3D Cartesian for your own research, please cite
at least one of the following articles:
\begin{description}
\item [{Numerical simulations in general}] ~\\
 Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
or \citet{KoVi98}. Additional aspects of adjoint simulations are
described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \citet{MaKoBlLe08},
and excellent scaling up to 150,000 processor cores is shown for instance
in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
\item [{GPU computing}] ~\\
 Computing on GPU graphics cards for acoustic or seismic wave propagation
applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
\end{description}
\noindent If you use this new version, which has non blocking
MPI for much better performance for medium or large runs, please cite
at least one of these six articles, in which results of non blocking
MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\

\noindent If you work on geophysical applications, you may be interested
in citing some of these application articles as well, among others:
\begin{description}
\item [{Southern California simulations}] ~\\
 \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.

If you use the 3D southern California model, please cite \citet{SuSh03}
(Los Angeles model), \citet{lovelyetal06} (Salton Trough), and \citet{hauksson2000}
(southern California). The Moho map was determined by \citet{zhu&kanamori2000}.
The 1D SoCal model was developed by \citet{DrHe90}.

\item [{Anisotropy}] ~\\
 \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
\item [{Attenuation}] ~\\
 \citet{SaKoTr10,KoTr02a,KoTr99}.
\item [{Topography}] ~\\
 \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
\end{description}
The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.

\section{Support}

This material is based upon work supported by the USA National Science
Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
Reintegration Grant No. MIRG-CT-2005-017461. Any opinions, findings,
and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
USA National Science Foundation, CNRS, INRIA, ANR or the European
Marie Curie program.

\chapter{\label{cha:Getting-Started}Getting Started}

@@

\section{Using the ADIOS library for I/O}
Regular POSIX I/O can be problematic when dealing with large simulations one large
clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
to deal transparently take advantage of advanced paralllel file system features. To enable
ADIOS, the following steps should be done:
\begin{enumerate}
\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
Make sure that your environment variables reference it.
\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
The default values probably suit most cases.
\item Configure using the \texttt{--with-adios} flag.
\end{enumerate}
ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{\label{cha:Mesh-Generation}Mesh Generation}

The first step in running a spectral-element simulation consists of
constructing a high-quality mesh for the region under consideration.
We provide two possibilities to do so: (1) relying on the external,
hexahedral mesher CUBIT, or (2) using the provided, internal mesher
\texttt{xmeshfem3D}. In the following, we explain these two approaches.

\section{\label{cha:Running-the-Mesher-CUBIT}Meshing with \texttt{CUBIT}}

CUBIT is a meshing tool suite for the creation of finite-element meshes
for arbitrarily shaped models. It has been developed and maintained
at Sandia National Laboratories and can be purchased for a small academic
institutional fee at \url{http://cubit.sandia.gov}. Our experience
showed that using CUBIT greatly facilitates and speeds up the generation
and preparation of hexahedral, conforming meshes for a variety of
geophysical models with increasing complexity.

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.6\textwidth]{figures/mount-cubit}
\par\end{centering}

\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
shown in the main display consists of a hexahedral discretization
of a single volume with topography.}

\label{fig:mount.cubit}
\end{figure}

\noindent
The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

\subsection{Creating the Mesh with CUBIT}

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
 \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
\par\end{centering}

\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}

\label{fig:examples.cubit}
\end{figure}

In each example subdirectory you will find a \texttt{README} file,
which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package by
contacting the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

In some cases, to re-create the meshes for the examples given, just type "claro ./create\_mesh.py" or similar
from the command line ("claro" is the command to run CUBIT from the command line).

IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
in particular, please make sure you correctly set up the Python paths as indicated in that file.

\subsection{Exporting the Mesh with \texttt{run\_cubit2specfem3d.py} }

\label{subsec:Exporting-the-Mesh}

Once the geometric model volumes in CUBIT are meshed, you prepare
the model for exportation with the definition of material blocks and
boundary surfaces. Thus, prior to exporting the mesh, you need to
define blocks specifying the materials and absorbing boundaries in
CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
if the mesh meets some conditions or manually, following the block
convention:
\begin{description}
\item [{material\_name}] Each material should have a specific block defined
by a unique name. The name convention of the material is to start
with either 'elastic' or 'acoustic'. It must be then followed by a
unique identifier, e.g. 'elastic 1', 'elastic 2', etc. The additional
attributes to the block define the material description.

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. Please note that
your Vp- and Vs-speeds are given for a reference frequency. To change
this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
in the main constants file \texttt{constants.h} found in subdirectory
\texttt{src/shared/}.
\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
in case an anisotropic simulation should be conducted. See the file
\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
for an implementation of the anisotropic models. In case no anisotropy
is available, it can be set to zero.
\end{description}

Note that this material block has to be defined using all the volumes
which belong to this elastic material. For volumes belonging to another,
different material, you will need to define a new material block.

For an acoustic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{0}] S-wave speed of the material is ignored.
\item [{rho}] density of the material (given in kg/m$^{3}$).
\end{description}
\item [{face\_topo}] Block definition for the surface which defines the
free surface (which can have topography). The name of this block must
be 'face\_topo', the block has to be defined using all the surfaces
which constitute the complete free surface of the model.
\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmin.
\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmax.
\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymin.
\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymax.
\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
boundaries, one block for each surface with z=bottom.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-surface} \includegraphics[width=0.45\textwidth]{figures/mount-abs}
\par\end{centering}

\caption{Example of the block definitions for the free surface 'face\_topo'
(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
(right) in CUBIT.}

\label{fig:mount.abs}
\end{figure}

Optionally, instead of specifying for each surface at the boundaries
a single block like mentioned above, you can also specify a single
block for all boundary surfaces and name it as one of the absorbing
blocks above, e.g. 'face\_abs\_xmin'.

After the block definitions are done, you export the mesh using the
script \texttt{cubit2specfem3d.py} provided in each of the example
directories (linked to the common script \texttt{CUBIT\_GEOCUBIT/cubit2specfem3d.py}).
If the export was successful, you should find the following
files in a subdirectory \texttt{MESH/}:
\begin{description}
\item [{absorbing\_cpml\_file}] \textbf{(only needed in case of C-PML absorbing
conditions)} Contains on the first line the total number of C-PML
spectral elements in the mesh, and then on the following line the
list of all these C-PML elements with two numbers per line: first
the spectral element number, and then a C-PML flag indicating to which
C-PML layer(s) that element belongs, according to the following convention:

\begin{itemize}
\item Flag = 1 : element belongs to a X CPML layer only (either in Xmin
or in Xmax),
\item Flag = 2 : element belongs to a Y CPML layer only (either in Ymin
or in Ymax),
\item Flag = 3 : element belongs to a Z CPML layer only (either in Zmin
or in Zmax),
\item Flag = 4 : element belongs to a X CPML layer and also to a Y CPML
layer,
\item Flag = 5 : element belongs to a X CPML layer and also to a Z CPML
layer,
\item Flag = 6 : element belongs to a Y CPML layer and also to a Z CPML
layer,
\item Flag = 7 : element belongs to a X, to a Y and to a Z CPML layer, i.e.,
it belongs to a CPML corner.
\end{itemize}

Note that it does not matter whether an element belongs to a Xmin
or to a Xmax CPML, the flag is the same in both cases; the same is
true for Ymin and Ymax, and also Zmin and Zmax.

When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
format, i.e., if you have existing \textquotedbl{}nodes\_coords\_file\textquotedbl{}
and \textquotedbl{}mesh\_file\textquotedbl{} files but do not know
how to assign CPML flags to them, we have created a small serial Fortran
program that will do that automatically for you, i.e., which will
create the \textquotedbl{}absorbing\_cpml\_file\textquotedbl{} for
you. That program is utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90,
and a small Makefile is provided in that directory (utils/CPML).

IMPORTANT: it is your responsibility to make sure that in the input
CUBIT (or similar) mesh that this code will read in SPECFEM3D format
from files \textquotedbl{}nodes\_coords\_file\textquotedbl{} and \textquotedbl{}mesh\_file\textquotedbl{}
you have created layers of elements that constitute a layer of constant
thickness aligned with the coordinate grid axes (X, Y and/or Z), so
that this code can assign CPML flags to them. This code does NOT check
that (because it cannot, in any easy way). The mesh inside these CPML
layers does not need to be structured nor regular, any non-structured
mesh is fine as long as it has flat PML inner and outer faces, parallel
to the axes, and thus of a constant thickness. The thickness can be
different for the X, Y and Z sides. But for X it must not vary, for
Y it must not vary, and for Z it must not vary. If you do not know
the exact thickness, you can use a slightly LARGER value in this code
(say 2\% to 5\% more) and this code will fix that and will adjust
it; never use a SMALLER value otherwise this code will miss some CPML
elements.

Note: in a future release we will remove the constraint of having
CPML layers aligned with the coordinate axes; we will allow for meshes
that are titled by any constant angle in the horizontal plane. However
this is not implemented yet.

Note: in the case of fluid-solid layers in contact, the fluid-solid
interface currently needs to be flat and horizontal inside the CPML
layer (i.e., bathymetry should become flat and horizontal when entering
the CPML); this (small) constraint will probably remain in the code
for a while because it makes fluid-solid matching inside the CPML
much easier.

\item [{materials\_file}] Contains the material associations for each element.
The format is:

\begin{lyxcode}
element\_ID~material\_ID
\end{lyxcode}

>>>>

\chapter{Getting Started}\label{cha:Getting-Started}

@@

\section{Using the ADIOS library for I/O}

Regular POSIX I/O can be problematic when dealing with large simulations one large
clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
to deal transparently take advantage of advanced paralllel file system features. To enable
ADIOS, the following steps should be done:
\begin{enumerate}
\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
Make sure that your environment variables reference it.
\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
The default values probably suit most cases.
\item Configure using the \texttt{--with-adios} flag.
\end{enumerate}
ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.

@@
